\chapter*{Acknowledgments}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

This thesis would not have been possible without the contributions, explicit or implicit, of many.
First and foremost, I would like to thank my advisors, Patrick Thiran and Matthias Grossglauser, for their mentorship.
Their creativity, rigour, curiosity, and generosity were a true inspiration.
Thank you, Patrick and Matthias, for believing in me when I did not.

I had the honor of having Bob West, Scott Hale, Rayid Ghani, and Karl Aberer as members of my jury committee.
Thank you all for your time and your excellent comments and suggestions.
I really enjoyed discussing the five and half years of my work with you.

I have been fortunate to be surrounded by a fantastic staff at the lab.
Holly Cogliati-Bauereis provided invaluable help by proof-reading and improving my manuscripts.
I became a much better writer thanks to her feedback (and her crisp humour!).
Patricia Hjelt and Angela Devenoge assisted me with the administrative tasks---always with utmost kindness.
I really enjoyed our personal discussions and their genuine interest in other people's lives.
Marc-André Lüthi and Yves Lopes helped me with our IT infrastructure, which made tuning hyperparameters of predictive models an (almost) pleasant task.
I thank them all.

I am very grateful to two special people, whom I consider my academic role models.
Their guidance was invaluable to my obtention of this degree.
Vincent Etter, I thank you for being an influential variable in my decision to begin my Ph.D.\ research.
Since then, you have always been available to discuss my ideas and me give me honest feedback.
Lucas Maystre, I thank you for your kindness, patience, positivity, and knowledge.
I learned so much from you about so many aspects of doing research.

I also want to express my gratitude to my academic collaborators, in particular to Batuhan Yardim, Alexander Immer, and Aswin Suresh for being the best co-authors I could ever hope for.
Working with all of you was a pure joy and very inspiring.
I thank Roy Gava and Steven Eichenberger for co-organizing the AI \& Democracy conference, Marlene Kammerer and Paula Castro for our project about international climate negotiations, Corinne Straub for building the API that made Predikon a reality, Jérôme Payet, Alexis Barrou, Edouard Cattin, Blanche Dalimier, Valentin Quelquejay-Leclere, and Robin Zbinden for our joint efforts to create Climpact.
I am also very grateful to Augustin Chaintreau and Ana-Andreea Stoica who welcomed me with kindness and generosity at Columbia University during the Summers of 2017 and 2018.

I was delighted to work with a group of smart and fun people who gave supportive advice throughout my time in the INDY Lab.
Thank you
Alexandre,
Arnout,
Brunella,
Christina,
Daniyar,
Ehsan,
Emti,
Farnood,
Greg,
Julien,
Lars,
Ljudmila,
Mahsa,
Mladen,
Mohamed,
Sébastien,
A very special thank you goes to William for being my fellow traveller during the sinuous journey of our theses.
You were a motivation to come to the lab every day and were always there to support me when times were tough.
I will keep our coffees in the lounge, beers at Sat', lunches at the food trucks, and discussions about work and life as some of the best memories of these years.

I am also grateful to everyone at Swiss Youth for Climate during the four years I was part of this amazing adventure.
Making the world a better place is an arduous task, but being surrounded by incredible people makes it fun.
This experience also taught me invaluable skills that I would never have learned behind my computer in the lab.
I also extend thank yous to everyone involved in the Divest campaign at EPFL.
I am proud that our efforts contributed to making our university more environmentally friendly.

Finally, I express my deep gratitude to my family.
Thank you for your trust, understanding, and support.
I am grateful, most of all, to Lydie-Line.
Your empathy, sensitivity, humour, generosity, and love make everything in life enjoyable and possible---even a Ph.D.\ degree.

\bigskip

\noindent\textit{Lausanne, May 31, 2021}
\hfill V.~K.
